---
title: "Environmental Justice Data Exploration"
author: "Sofia Rodas"
date: October 6, 2025
format: html
---

## Introduction
This quarto document explores environmental justice issues in NorthEast Los Angeles (NELA), specifically Eagle Rock, Highland Park, Glassell Park, Cypress Park, Atwater Village, Montecito Heights, Lincoln Heights, El Sereno, and Mount Washington. There is some dispute about the official Los Anegeles neighborhoods that make up NELA so Matsuoka & Urquiza's NELA definition was utilized for this analysis. NELA is ethnically diverse making it an interesting sector to analyze. 

Matsuoka, Martha & Urquiza, John. (2021). Building community knowledge, resilience and resistance through research. GeoJournal. 87. 10.1007/s10708-021-10422-5. 

#### Adding the necessary libraries:
```{r}
library(tidyverse)
library(sf)
library(here)
library(tmap)
```
## Import data
```{r}
# Read in geodatabase of EJScreen data at the Census Block Group level
ejscreen <- sf::st_read(here::here("data", "ejscreen","EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb")) 
 
# Read in Los Angeles neighnorhood shape file
la_neighborhood <- sf::st_read(here::here("LA_Times_Neighborhood_Boundaries-shp", "8494cd42-db48-4af1-a215-a2c8f61e96a22020328-1-621do0.x5yiu.shp"))
```
## Subset Data
###### Changing column names to lower_snake
```{r}
ejscreen <- janitor::clean_names(ejscreen)
```
I'm interested in looking at cancer rates and how they correlate to the percentage of low income individuals in NorthEast Los Angeles. 

I used a Los Angeles neighborhoods shape file to select for the region that from 
https://geohub.lacity.org/datasets/d6c55385a0e749519f238b77135eafac/explore?location=34.269713%2C-117.262418%2C8.80&showTable=true 

##### Filter for Los Angeles County 
```{r}
los_angeles <- ejscreen |> 
  filter(cnty_name == "Los Angeles County")
```

#### Filter for NELA neighborhoods
```{r}
la_neighborhood <- la_neighborhood |> 
  filter(name == "Eagle Rock" | name == "Highland Park" | name == "Mount Washington" | name == "Glassell Park" | name == "Cypress Park"| name == "Montecito Heights"| name == "Atwater Village"| name == "Lincoln Heights"| name == "El Sereno"  )
```

# Visualize Data
### Map: People of Color Percentage in NorthEast Los Angeles 
```{r}
 poc_percentiles <- tm_shape(los_angeles)+
  tm_polygons(fill = "peopcolorpct", 
              fill.scale = tm_scale(values = "brewer.YlOrBr"),
              fill.legend = tm_legend(title = "Percentile People of Color")) +
tm_shape(la_neighborhood, is.main = TRUE) + # is.main to set the size of map
  tm_borders() + # Use tm_borders so the layer under is visible
  tm_text("name", # Name helps distinguish what part of LA we're looking at
          shadow = TRUE, 
          bgcol = "white") +
  tm_title(text = "NorthEast Los Angeles: People of Color Percentiles",
           frame = TRUE)+
  tm_compass(position = c("right", "top"), 
             bg = TRUE,
             bg.color = "white",
             frame = TRUE)+
  tm_scale_bar(position = c("left", "bottom"),
               bg = TRUE,
               bg.col = "white") 

# Save map and define the size
tmap_save(poc_percentiles, 
          here::here("maps", "poc_percentiles.png"), 
          height = 8, 
          width = 14)

# Show map in the quarto doc 
print(poc_percentiles)
```
### Map: Air toxics cancer risk, an Environmental Justice (EJ) Index in NorthEast Los Angeles 
```{r}
air_cancer_risk <- tm_shape(los_angeles)+
  tm_polygons(fill = "d2_cancer",
              fill.scale = tm_scale(values = "brewer.YlOrBr"),
              fill.legend = tm_legend(title = "Air Toxins Cancer Risk EJ Index Percentile")) +
tm_shape(la_neighborhood, is.main = TRUE) + #is.main to set the size of the map
  tm_borders() + # Use tm_borders so the layer under is visible
  tm_text("name", # Name helps distinguish what part of LA we're looking at
          shadow = TRUE, 
          bgcol = "white") +
  tm_title(text = "NorthEast Los Angeles: Air Toxins Cancer Risk EJ Index",
           frame = TRUE)+
  tm_compass(position = c("right", "top"), 
             bg = TRUE,
             bg.color = "white",
             frame = TRUE) +
  tm_scale_bar(position = c("left", "bottom"),
               bg = TRUE,
               bg.col = "white")

# Save map and define the size
tmap_save(air_cancer_risk, 
          here::here("maps", "air_cancer_risk.png"), 
          height = 8, 
          width = 14)

# Show map in the quarto doc 
print(air_cancer_risk)
```

# Map Interpretation
The first graph above shows people of color percentage for each block. The opacity of the orange darkens as the percentage of people of color increases. The second graph shows air toxins cancer risk environmental justice index. The air toxins cancer risk environmental justice index is based on the estimated risk of inhaling carcinogens in the air. There is an obvious correlation showing that the more people of color live in an area, the more likely they are to have a high score for air toxins cancer risk environmental justice risk index. 

# References


Matsuoka, Martha & Urquiza, John. (2021). Building community knowledge, resilience and resistance through research. GeoJournal. 87. 10.1007/s10708-021-10422-5.