---
title: "Environmental Justice Data Exploration"
author: "Sofia Rodas"
date: October 6, 2025
format: docx
---

## Introduction
This quarto document explores environmental justice issues in NorthEast Los Angeles (NELA), specifically Eagle Rock, Highland Park, Glassell Park, Cypress Park, Atwater Village, Montecito Heights, Lincoln Heights, El Sereno, and Mount Washington. There is some dispute about the official Los Anegeles' neighborhoods that make up NELA so [Matsuoka & Urquiza's](https://www.researchgate.net/publication/351828642_Building_community_knowledge_resilience_and_resistance_through_research) NELA definition was utilized for this analysis(2021). NELA is an interesting case study to analyze as it is ethnically diverse and can act as a representation for what may be occuring across the United States. 

 

#### Adding the necessary libraries:
```{r}
library(tidyverse)
library(sf)
library(here)
library(tmap)
```
## Import data
```{r}
# Read in geodatabase of EJScreen data at the Census Block Group level
ejscreen <- sf::st_read(here::here("data",
                                   "ejscreen",
                                   "EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb")) 
 
# Read in Los Angeles neighnorhood shape file
la_neighborhood <- sf::st_read(here::here("LA_Times_Neighborhood_Boundaries-shp", 
                                          "8494cd42-db48-4af1-a215-a2c8f61e96a22020328-1-621do0.x5yiu.shp"))
```
## Subset Data
###### Changing column names to lower_snake
```{r}
ejscreen <- janitor::clean_names(ejscreen)
```

#### Filter for Los Angeles County 
```{r}
los_angeles <- ejscreen |> 
  filter(cnty_name == "Los Angeles County")
```

#### Filter for NELA neighborhoods
```{r}
la_neighborhood <- la_neighborhood |> 
  filter(name == "Eagle Rock" |
           name == "Highland Park" |
           name == "Mount Washington" | 
           name == "Glassell Park" | 
           name == "Cypress Park"| 
           name == "Montecito Heights"| 
           name == "Atwater Village"| 
           name == "Lincoln Heights"| 
           name == "El Sereno"  )
```

## Data Manipulation 
#### Changing the measure from 0-1 to 0-100
```{r}
los_angeles <- los_angeles |> 
  mutate(poc_ile = (peopcolorpct * 100))
```


# Visualize Data
### Map: People of Color Percentile in NorthEast Los Angeles 
```{r}
 poc_percentiles <- tm_shape(los_angeles)+
  tm_polygons(fill = "poc_ile", 
              fill.scale = tm_scale(values = "brewer.YlOrBr"),
              fill.legend = tm_legend(title = "People of Color Percentile")) +
tm_shape(la_neighborhood, is.main = TRUE) + # is.main to set the size of map
  tm_borders() + # Use tm_borders so the layer under is visible
  tm_text("name", # Name helps distinguish what part of LA is being mapped
          shadow = TRUE, 
          bgcol = "white") +
  tm_title(text = "NorthEast Los Angeles: People of Color Percentiles",
           frame = TRUE)+
  tm_compass(position = c("right", "top"), 
             bg = TRUE,
             bg.color = "white",
             frame = TRUE)+
  tm_scale_bar(position = c("left", "bottom"),
               bg = TRUE,
               bg.col = "white") 

# Save map and define the size
tmap_save(poc_percentiles, 
          here::here("maps", "poc_percentiles.png"),
          height = 8, 
          width = 14)

# Show map in the quarto doc 
print(poc_percentiles)
```
### Map: Air Toxics Cancer Risk Percentile, an Environmental Justice (EJ) Index in NorthEast Los Angeles 
```{r}
air_cancer_risk <- tm_shape(los_angeles)+
  tm_polygons(fill = "d2_cancer",
              fill.scale = tm_scale(values = "brewer.YlOrBr"),
              fill.legend = tm_legend(title = "Air Toxins Cancer Risk EJ Index Percentile")) +
tm_shape(la_neighborhood, is.main = TRUE) + #is.main to set the size of the map
  tm_borders() + # Use tm_borders so the layer under is visible
  tm_text("name", # Name helps distinguish what part of LA is being mapped
          shadow = TRUE, 
          bgcol = "white") +
  tm_title(text = "NorthEast Los Angeles: Air Toxins Cancer Risk EJ Index Percentiles",
           frame = TRUE)+
  tm_compass(position = c("right", "top"), 
             bg = TRUE,
             bg.color = "white",
             frame = TRUE) +
  tm_scale_bar(position = c("left", "bottom"),
               bg = TRUE,
               bg.col = "white")

# Save map and define the size
tmap_save(air_cancer_risk, 
          here::here("maps", "air_cancer_risk.png"), 
          height = 8, 
          width = 14)

# Show map in the quarto doc 
print(air_cancer_risk)
```

# Map Interpretation
The first graph above shows people of color percentile for each Census block. The second graph shows percentile of air toxics cancer risk environmental justice index. The opacity of the orange darkens as the percentiles increase. The air toxics cancer risk environmental justice index is based on the estimated risk of inhaling carcinogens in the ambient air. There is an obvious correlation showing that the higher the percentile of people of color live in an area, the more likely a community is to have a high percentile for air toxics cancer risk environmental justice index. In other words, people of color are exposed to cancer risks from their ambient environment at a higher rate.

Though disheartening, the finding that people of color live where environmental pollution health risks are greater, provides knowledge of the issue and is the first step to making policy change that can help offset injust risks. Air toxics is one of the many environmental justice indices. Further analysis of the distribution of risks ranging from toxic wastewater discharge to lead paint in homes can help show accumulated environmental risk for people of color and show the immediate need for intervention to reduce negative outcomes for marginalized communities.
 

# References

- LA Times. (2016). LA Times Neighborhood Boundaries. Retrieved from [URL](https://geohub.lacity.org/datasets/d6c55385a0e749519f238b77135eafac_0/about)
- [Matsuoka, Martha & Urquiza, John. (2021). Building community knowledge, resilience and resistance through research. GeoJournal. 87. 10.1007/s10708-021-10422-5.](https://www.researchgate.net/publication/351828642_Building_community_knowledge_resilience_and_resistance_through_research)
- [U.S. Environmental Protection Agency (EPA), 2023. EJScreen Technical Documentation.](https://drive.google.com/file/d/1nG6Nj1bXfzQFOVMO8Km3eNy4SWu1YcIQ/view)