---
title: "Environmental Justice Data"
format: html
---
Adding the necessary libraries:
```{r}
library(tidyverse)
library(sf)
library(here)
library(tmap)
```
Reading in the data:
```{r}
# Read in geodatabase of EJScreen data at the Census Block Group level
ejscreen <- sf::st_read(here::here("data", "ejscreen","EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb")) 
 
# Read in Los Angeles zip code data
la_zip <- sf::st_read(here::here("Zip_Codes_(LA_County)", "Zip_Codes_(LA_County).shp"))

# Read in Los Angeles neighnorhood shape file
la_neighborhood <- sf::st_read(here::here("LA_Times_Neighborhood_Boundaries-shp", "8494cd42-db48-4af1-a215-a2c8f61e96a22020328-1-621do0.x5yiu.shp"))
```
Changing column names to lower_snake
```{r}
ejscreen <- janitor::clean_names(ejscreen)

la_zip <- janitor::clean_names(la_zip)
```
I'm interested in looking at cancer rates and how they correlate to the percentage of low income individuals in NorthEast Los Angeles. To concentrate the map shape on the desired location, I downloaded a LA County zip code shape file. Link https://geohub.lacity.org/datasets/lahub::zip-codes-la-county/about 

I uploaded Los Angeles neighborhoods from 
https://geohub.lacity.org/datasets/d6c55385a0e749519f238b77135eafac/explore?location=34.269713%2C-117.262418%2C8.80&showTable=true 

Filtering for Los Angeles County 
```{r}
los_angeles <- ejscreen |> 
  filter(cnty_name == "Los Angeles County")
```

Filtering for Eagle Rock and Highland Park
```{r}
la_zip <- la_zip |> 
  filter(zip == 90041 | zip == 90042 | zip == 90065 | zip == 90026)
  
```

Filtering for Eagle Rock, Highland Park, Mount Washington, and Glasell Park 
```{r}
la_neighborhood <- la_neighborhood |> 
  filter(name == "Eagle Rock" | name == "Highland Park" | name == "Mount Washington" | name == "Glassell Park" )
```


```{r}
tm_shape(la_neighborhood)+
  tm_polygons() +
tm_shape(los_angeles)+
  tm_polygons(fill = "lifeexppct",
              fill.scale = tm_scale(values = "brewer.YlOrBr")) 
  
  
```

```{r}
tm_shape(la_neighborhood)+
  tm_polygons() +
tm_shape(los_angeles)+
  tm_polygons(fill = "pre1960pct",
              fill.scale = tm_scale(values = "brewer.YlOrBr")) 
```

```{r}
# filter to a state you are interested in
california <- ejscreen |> 
  dplyr::filter(ST_ABBREV == "CA") 

# filter to a county you are interested in
santa_barbara <- ejscreen |> 
  dplyr::filter(CNTY_NAME %in% c("Santa Barbara County"))

# find the average values for all variables within counties
california_counties <- aggregate(california, by = list(california$CNTY_NAME), FUN = mean)
```

